QPLIB: A Library of Quadratic Programming Instances

### 2017-01-23 methods.tex
start to add references for solvers

Typically, bounds in the optimization direction are computed by means of convex relaxations \cite{mccormick,couenne}, which replace nonconvex terms $t(x)$ with linearization variables $\hat{t}$, and then replace the corresponding defining constraints $\hat{t} = t(x)$ by means of lower and upper (respectively, convex and concave) bounding functions $\hat{t} \ge \underline{t}(x)$ and $\hat{t} \le\bar{t}(x)$. This is actually where finite (and tight) bounds on the variables are crucial, which differentiate also in practice the bounded case from the unbounded one. Different strategies are used when the nonconvexities are only quadratic \cite{glomiqo,bliek}.

@@

\paragraph{Cutting plane approaches.}
%
Cutting plane approaches construct and iteratively improve a MILP (LIL) relaxation of the problem \cite{quesada,westerlund}. The cutting planes for the MILP are generated by linearization (first-order Taylor approximation) of the nonlinearities. If the latter are convex, the MILP provides a valid lower bound for the problem. Additionally, incomplete methods can be used to provide local solutions. Therefor, these methods are complete on CGC if a complete method is used to solve the MILP. The latter is typically based on BB, which is therefore a crucial technique also for this class of approaches.

Solvers in Table \ref{t:solvers} that implement complete cutting plane methods for CGC are \alphaecp \cite{westerlund}, \bonmin (in the algorithmic mode B-OA) \cite{BBCCGLLLMSW} and \dicopt \cite{quesada}. \todo{here we give references to the solvers, above we don't}

>>>>

Typically, bounds in the optimization direction are computed by means of convex relaxations \cite{mccormick,BeLeLiMaWa08}, which replace nonconvex terms $t(x)$ with linearization variables $\hat{t}$, and then replace the corresponding defining constraints $\hat{t} = t(x)$ by means of lower and upper (respectively, convex and concave) bounding functions $\hat{t} \ge \underline{t}(x)$ and $\hat{t} \le\bar{t}(x)$. This is actually where finite (and tight) bounds on the variables are crucial, which differentiate also in practice the bounded case from the unbounded one. Different strategies are used when the nonconvexities are only quadratic \cite{MiFl13,bliek}.

@@

\paragraph{Cutting plane approaches.}
%
Cutting plane approaches construct and iteratively improve a MILP (LIL) relaxation of the problem \cite{DuGr86,WePoe02}. The cutting planes for the MILP are generated by linearization (first-order Taylor approximation) of the nonlinearities. If the latter are convex, the MILP provides a valid lower bound for the problem. Additionally, incomplete methods can be used to provide local solutions. Therefor, these methods are complete on CGC if a complete method is used to solve the MILP. The latter is typically based on BB, which is therefore a crucial technique also for this class of approaches.

Solvers in Table \ref{t:solvers} that implement complete cutting plane methods for CGC are \alphaecp, \bonmin (in the algorithmic mode B-OA), and \dicopt.

### 2017-01-23 solvers.tex
make intro part to solution methods independent of set of solvers
start to add references for solvers

When compiling QPLIB, we have worked with QP solvers that come with the GAMS distribution\footnote{\url{https://www.gams.com}}.
A list of these solvers is provided by Table \ref{t:solvers}.

We mark a pair (solver, problem) with ``I'' if the solver accepts the problem in input but it is an incomplete solver for the problem, with ``A'' if it is asymptotically complete, with ``C'' if it is complete, and leave it blank if the solver won't accept the problem in input.
When a solver implements several algorithms, we have chosen, for each problem class, the algorithm that potentially provides the ``strongest'' results (``C'' $>$ ``A'' $>$ ``I'' $>$ blank).

\begin{table}
{
 \centering                
 \scriptsize                
 \setlength{\tabcolsep}{3pt}                
% \renewcommand \arraystretch{1}

\begin{tabular}{lccccccc}
\toprule  
             & CGL & QGL & CGC & QGQ & CCC & QCQ \\
\hline
\alphaecp    &  C  &  I  &  C  &  I  &  C  &  I  \\
\antigone    &  C  &  C  &  C  &  C  &  C  &  C  \\
\baron       &  C  &  C  &  C  &  C  &  C  &  C  \\
\bonmin      &  C  &  I  &  C  &  I  &  C  &  I  \\
\conopt      &     &     &     &     &  C  &  I  \\
\couenne     &  C  &  C  &  C  &  C  &  C  &  C  \\
\cplex       &  C  &  C  &  C  &     &  C  &     \\
\dicopt      &  C  &  I  &  C  &  I  &  C  &  I  \\
\gurobi      &  C  &     &  C  &     &  C  &     \\
\ipopt       &     &     &     &     &  C  &  I  \\
\knitro      &  C  &  I  &  C  &  I  &  C  &  A  \\
\lindo       &  C  &  C  &  C  &  C  &  C  &  C  \\
\lgo         &     &     &     &     &  A  &  A  \\
\minos       &     &     &     &     &  C  &  I  \\
\mosek       &  C  &     &  C  &     &  C  &     \\
\msnlp       &     &     &     &     &  C  &  A  \\
\oqnlp       &  A  &  A  &  A  &  A  &  C  &  A  \\
\sbb         &  C  &  I  &  C  &  I  &  C  &  I  \\
\scip        &  C  &  C  &  C  &  C  &  C  &  C  \\
\snopt       &     &     &     &     &  C  &  I  \\
\xpress      &  C  &     &  C  &     &  C  &     \\
\hline
\end{tabular}                 
\caption{Families of QP problems that can be tackled by each solver} \label{t:solvers}
}

@@

>>>>

When compiling QPLIB, we have worked with the QP solvers that come with the GAMS distribution\footnote{\url{https://www.gams.com}}. 
Table \ref{t:solvers} provides a list of these solvers, together with a classification of their algorithm, and references.
For more details on the solvers, we refer to the given references, solver manuals, and the survey \cite{BuVi10}.
In the table, we mark a pair (solver, problem) with ``I'' if the solver accepts the problem in input but it is an incomplete solver for the problem, with ``A'' if it is asymptotically complete, with ``C'' if it is complete, and leave it blank if the solver won't accept the problem in input.
When a solver implements several algorithms, we have chosen, for each problem class, the algorithm that potentially provides the ``strongest'' results (``C'' $>$ ``A'' $>$ ``I'' $>$ blank).

\begin{table}
 \centering
 \scriptsize
 \setlength{\tabcolsep}{3pt}

\begin{tabular}{llccccccc}
\toprule
             & & CGL & QGL & CGC & QGQ & CCC & QCQ \\
\midrule
\alphaecp    & \cite{WeLu03,WePoe02} &  C  &  I  &  C  &  I  &  C  &  I  \\
\antigone    & \cite{MiFl13,MiFl14} &  C  &  C  &  C  &  C  &  C  &  C  \\
\baron       & \cite{TaSa02,TaSa04,TaSa05} &  C  &  C  &  C  &  C  &  C  &  C  \\
\bonmin      & \cite{BBCCGLLLMSW} &  C  &  I  &  C  &  I  &  C  &  I  \\
\conopt      &  &     &     &     &     &  C  &  I  \\
\couenne     & \cite{BeLeLiMaWa08} &  C  &  C  &  C  &  C  &  C  &  C  \\
\cplex       & &  C  &  C  &  C  &     &  C  &     \\
\dicopt      & \cite{DuGr86,KoGr89,ViGr90} &  C  &  I  &  C  &  I  &  C  &  I  \\
\gurobi      & \cite{Ro12} &  C  &     &  C  &     &  C  &     \\
\ipopt       & &     &     &     &     &  C  &  I  \\
\knitro      & \cite{ByNoWa06} &  C  &  I  &  C  &  I  &  C  &  A  \\
\lindo       & \cite{LiSc09} &  C  &  C  &  C  &  C  &  C  &  C  \\
\lgo         & &     &     &     &     &  A  &  A  \\
\minos       & &     &     &     &     &  C  &  I  \\
\mosek       & &  C  &     &  C  &     &  C  &     \\
\msnlp       & &     &     &     &     &  C  &  A  \\
\oqnlp       & \cite{ULPGKM02} &  A  &  A  &  A  &  A  &  C  &  A  \\
\sbb         &  &  C  &  I  &  C  &  I  &  C  &  I  \\
\scip        & \cite{Ach09,VigerskeGleixner2016} &  C  &  C  &  C  &  C  &  C  &  C  \\
\snopt       & &     &     &     &     &  C  &  I  \\
\xpress      & &  C  &     &  C  &     &  C  &     \\
\bottomrule
\end{tabular}
\caption{Families of QP problems that can be tackled by each solver} \label{t:solvers}

@@

% antigone/glomiqo: sBB
% baron: sBB
% couenne: sBB
% knitro: IPM with CG, SLQP, convexBB + outerapproxBB
% lindo/lindoglobal: sBB
% scip: sBB
% msnlp/oqnlp: multistart
% alphaecp: cutting plane for pseudoconvex MINLP
% bonmin: convexBB
% dicopt: alternating projection NLP/MILP
% sbb: convexBB
% conopt: GRG (active set -- like SQP)
% ipopt: IPM w/ line search filter
% lgo: Lipschitz BB, random search, multistart (default) [cassato?]
% minos: active set method
% snopt: active set SQP
% xpress: convexBB
% gurobi: convexBB
% cplex: sBB
